\documentclass[a4wide,10pt]{article}
%\usepackage{a4wide}
\usepackage[applemac,utf8]{inputenc}
\usepackage[danish]{babel}
\usepackage[T1]{fontenc}
\usepackage{pdfsync}
\usepackage{amsmath,amssymb,amsfonts} 
\usepackage[pdftex]{graphicx}
\usepackage{wrapfig}
\usepackage{color}
\usepackage[small,bf]{caption}

\begin{document}
\title{COS3 lektion 8}
\author{Nis Sarup}
\date{\today}
\maketitle


\addtocounter{section}{7}
\section{Memory Management} % (fold)
\label{sec:memory_management}
\subsection{Background} % (fold)
\label{sub:background}
\begin{itemize}
	\item Registers are fast
	\item Memory is not fast
	\item Each process needs a separate memory space set by a base and limiter register
	\item Attempting to write to memory outside the process own space is denied
	\item An address for a memory segment from the CPU: logical address
	\item An address for a memory segment from the memory unit: physical address
	\item Dynamic loading: ONly load what's needed
\end{itemize}
% subsection background (end)

\subsection{Swapping} % (fold)
\label{sub:swapping}
\begin{itemize}
	\item Inactive processes are not kept in memory, but relocated to other storage. Fast disk, etc.
	\item Swapping is rather slow.
\end{itemize}
% subsection swapping (end)

\subsection{Contiguous Memory Allocation} % (fold)
\label{sub:contiguous_memory_allocation}
\begin{itemize}
	\item Each process is contained in a single contiguous section of memory.
	\item Multiple partition method: Memory is partitioned into fixed width segments. Each process gets a segment.
	\item Variable Partition: A table of "holes" in the memory are maintained. A new process gets a ssigned to a matching hole.
	\begin{itemize}
		\item First Fit: First hole that's big enough.
		\item Best Fit: Smallest hole that's big enough.
		\item Worst Fit: Largest hole is selected.
	\end{itemize}
	\item 50-percent rule: Given N allocated blocks, another 0.5 N blocks is lost to fragmentation.
	\item Compaction: Shuffle memory contents to make one big hole.
\end{itemize}
% subsection contiguous_memory_allocation (end)

\subsection{Paging} % (fold)
\label{sub:paging}
\begin{itemize}
	\item Paging permits the physical address space of a process to be non-contiguous.
	\item Paging can be implemented in hardware and in software.
	\item Logical memory is split into pages
	\item Physical memory is split into frames
	\item The page table holds which page resides on which frame
	\item Pages can be shared between processes.
\end{itemize}
% subsection paging (end)

\subsection{Structure of the Page Table} % (fold)
\label{sub:structure_of_the_page_table}
\begin{itemize}
	\item Large Page tables can be page tabled.
\end{itemize}
% subsection structure_of_the_page_table (end)
% section memory_management (end)
\end{document}